\pkg{cv}: An \proglang{R} Package for Cross-Validation of Regression Models

### 2024-01-13 scratch/notes_john/JSS-article/JSS-article-reduced-2.Rmd
further edits to JSS draft

The \code{Auto} appeared in a preliminary example in Section \ref{preliminary-example-polynomial-regression}, where we employed CV to inform the selection of the degree of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set (Figure \ref{fig:Auto-explore}, produced by the \code{scatterplotMatrix()} function in the \pkg{car} package):
```{r Auto-explore}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Scatterplot matrix for the numeric variables in the Auto data"
Auto$cylinders <- factor(Auto$cylinders,
                         labels=c("3.4", "3.4", "5.6", "5.6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels=c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique=TRUE)
Auto$name <- NULL

@@

[^year]: Of course, making the decision to treat \code{year} as a factor on this basis could be construed as cheating in the current context, which illustrates the difficulty of automating the whole model-selection process. It's rarely desirable, in our opinion, to forgo exploration of the data to ensure the purity of model validation. We believe, however, that it's still useful to automate as much of the process as we can to obtain a more realistic, if still biased, estimate of the predictive power of a model.

@@

* \code{data}, the data set to which the model is fit.
* \code{seed}, an optional seed for \proglang{R}'s pseudo-random-number generator; as for \code{cv()}, if the seed isn't supplied by the user, then a seed is randomly selected and saved.
* additional arguments required by the model-selection function, here the starting \code{model} for predictor selection; the names of the variables, \code{predictors} and \code{response}, that are candidates for transformation; and \code{AIC=FALSE}, which specifies use of the BIC for model selection. The starting model, \code{m.auto}, is the working model fit to the \code{Auto} data.

>>>>

The \code{Auto} data appeared in a preliminary example in Section \ref{preliminary-example-polynomial-regression}, where we employed CV to inform the selection of the degree of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set (Figure \ref{fig:Auto-explore}, produced by the \code{scatterplotMatrix()} function in the \pkg{car} package):
```{r Auto-explore}
#| out.width = "60%",
#| fig.height = 5,
#| fig.cap = "Scatterplot matrix for the numeric variables in the Auto data"
Auto$cylinders <- factor(Auto$cylinders,
                         labels=c("3.4", "3.4", "5.6", "5.6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels=c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique=TRUE)
Auto$name <- NULL

@@

[^year]: Making the decision to treat \code{year} as a factor on this basis could be construed as cheating in the current context, which illustrates the difficulty of automating the whole model-selection process. It's rarely desirable, in our opinion, to forgo exploration of the data to ensure the purity of model validation. We believe, however, that it's still useful to automate as much of the process as we can to obtain a more realistic, if still biased, estimate of the predictive power of a model.

@@

* \code{data}, the data set to which the model is fit.
* \code{seed}, an optional seed for \proglang{R}'s pseudo-random-number generator; as for \code{cv()}, if the seed isn't supplied by the user, then a seed is randomly selected and saved.
* Arguments required by the model-selection function: the starting \code{model} (\code{m.auto}) for predictor selection, the names of the variables---\code{predictors} and \code{response}---that are candidates for transformation, and \code{AIC=FALSE}, which specifies use of the BIC for model selection.

### 2024-01-13 scratch/notes_john/JSS-article/JSS-article-reduced-2.tex
further edits to JSS draft

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 392419
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.16505
bias-adjusted cross-validation criterion = 19.15554
95% CI for bias-adjusted CV criterion = (15.71577, 22.59531)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

The \code{Auto} appeared in a preliminary example in Section
\ref{preliminary-example-polynomial-regression}, where we employed CV to
inform the selection of the degree of a polynomial regression of
\code{mpg} on \code{horsepower}. Here, we consider more generally the
problem of predicting \code{mpg} from the other variables in the
\code{Auto} data. We begin with a bit of data management, and then
examine the pairwise relationships among the numeric variables in the
data set (Figure \ref{fig:Auto-explore}, produced by the
\code{scatterplotMatrix()} function in the \pkg{car} package):

@@

A comment before we proceed: \code{origin} is clearly categorical and so
converting it to a factor is natural, but we could imagine treating
\code{cylinders} and \code{year} as numeric predictors. There are,
however, only 5 distinct values of \code{cylinders} (ranging from 3 to
8), but cars with 3 or 5 cylinders are rare, and none of the cars has 7
cylinders. There are similarly only 13 distinct years between 1970 and
1982 in the data, and the relationship between \code{mpg} and
\code{year} is difficult to characterize.\footnote{Of course, making the
  decision to treat \code{year} as a factor on this basis could be
  construed as cheating in the current context, which illustrates the
  difficulty of automating the whole model-selection process. It's
  rarely desirable, in our opinion, to forgo exploration of the data to
  ensure the purity of model validation. We believe, however, that it's
  still useful to automate as much of the process as we can to obtain a
  more realistic, if still biased, estimate of the predictive power of a
  model.} It's apparent that most these variables are positively skewed
and that many of the pairwise relationships among them are nonlinear.

@@

\begin{itemize}
\tightlist
\item
  \code{data}, the data set to which the model is fit.
\item
  \code{seed}, an optional seed for \proglang{R}'s pseudo-random-number
  generator; as for \code{cv()}, if the seed isn't supplied by the user,
  then a seed is randomly selected and saved.
\item
  additional arguments required by the model-selection function, here
  the starting \code{model} for predictor selection; the names of the
  variables, \code{predictors} and \code{response}, that are candidates
  for transformation; and \code{AIC=FALSE}, which specifies use of the
  BIC for model selection. The starting model, \code{m.auto}, is the
  working model fit to the \code{Auto} data.
\end{itemize}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat){
  if (!all(y %in% c(0, 1))) stop("response values not all 0 or 1")
  if (any(yhat < 0) || any(yhat > 1)) stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13ca97218>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

>>>>

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 352413
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.42674
bias-adjusted cross-validation criterion = 19.40347
95% CI for bias-adjusted CV criterion = (15.89518, 22.91177)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

The \code{Auto} data appeared in a preliminary example in Section
\ref{preliminary-example-polynomial-regression}, where we employed CV to
inform the selection of the degree of a polynomial regression of
\code{mpg} on \code{horsepower}. Here, we consider more generally the
problem of predicting \code{mpg} from the other variables in the
\code{Auto} data. We begin with a bit of data management, and then
examine the pairwise relationships among the numeric variables in the
data set (Figure \ref{fig:Auto-explore}, produced by the
\code{scatterplotMatrix()} function in the \pkg{car} package):

@@

A comment before we proceed: \code{origin} is clearly categorical and so
converting it to a factor is natural, but we could imagine treating
\code{cylinders} and \code{year} as numeric predictors. There are,
however, only 5 distinct values of \code{cylinders} (ranging from 3 to
8), but cars with 3 or 5 cylinders are rare, and none of the cars has 7
cylinders. There are similarly only 13 distinct years between 1970 and
1982 in the data, and the relationship between \code{mpg} and
\code{year} is difficult to characterize.\footnote{Making the decision
  to treat \code{year} as a factor on this basis could be construed as
  cheating in the current context, which illustrates the difficulty of
  automating the whole model-selection process. It's rarely desirable,
  in our opinion, to forgo exploration of the data to ensure the purity
  of model validation. We believe, however, that it's still useful to
  automate as much of the process as we can to obtain a more realistic,
  if still biased, estimate of the predictive power of a model.} It's
apparent that most these variables are positively skewed and that many
of the pairwise relationships among them are nonlinear.

@@

\begin{itemize}
\tightlist
\item
  \code{data}, the data set to which the model is fit.
\item
  \code{seed}, an optional seed for \proglang{R}'s pseudo-random-number
  generator; as for \code{cv()}, if the seed isn't supplied by the user,
  then a seed is randomly selected and saved.
\item
  Arguments required by the model-selection function: the starting
  \code{model} (\code{m.auto}) for predictor selection, the names of the
  variables---\code{predictors} and \code{response}---that are
  candidates for transformation, and \code{AIC=FALSE}, which specifies
  use of the BIC for model selection.
\end{itemize}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat){
  if (!all(y %in% c(0, 1))) stop("response values not all 0 or 1")
  if (any(yhat < 0) || any(yhat > 1)) stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x128c4cd38>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}
